Contextuality and the Born rule in fuzzy quantum theories

### 2017-03-26 prop_letter.tex
prop_letter.tex Updaitng according to "APS Author Guide for REVTEX 4.1" until "V.7. Boxes"... I don't know how to handle \mbox so far...

\documentclass[english,reprint, aps, prl,superscriptaddress, showpacs,
showkeys]{revtex4-1} 
\usepackage{fullpage}
\usepackage[T1,T2A]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{babel}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{mathrsfs}
\usepackage{bbold}
\usepackage{wesa}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage{hyperref}

@@

\author{Andrew J Hanson}
\affiliation{School of Informatics and Computing, Indiana University, Bloomington,
IN 47405, USA}

\author{Gerardo Ortiz}
\affiliation{Department of Physics, Indiana University, Bloomington, IN 47405,
USA}

\author{Amr Sabry}
\affiliation{School of Informatics and Computing, Indiana University, Bloomington,
IN 47405, USA}

\author{Yu-Tsung Tai}
\affiliation{Department of Mathematics, Indiana University, Bloomington, IN 47405,
USA}
\affiliation{School of Informatics and Computing, Indiana University, Bloomington,
IN 47405, USA}

@@

\begin{abstract}
\begin{description}
\item [{Background}] In a world of bounded resource, any quantum measurement
can be only performed in finite precision. {\small \par}
\item [{Method}] We model the idea of the finite precision measurement
as finite number of intervals, and replace quantum real-valued probability
measures by interval-valued ones.{\small \par}
\item [{Results}] Quantum interval-valued probability measures provide
a unified framework to interpret Gleason's theorem and the Kochen-Specker
theorem as the limiting cases. For the cases between them, we found
some legitimate quantum interval-valued probability measures which
cannot be explained by the Born rule. While restricting on only commuting
projectors, the interval-valued probability measures become classical,
and these non-Born interval-valued probability measures disappear. {\small \par}
\item [{Conclusions}] Our research provides a new insight of finite precision
measurement and non-commutativity when formulating quantum theory.
Also, the Born rule may not be as nature as we used to believe in
the world of finite resource.{\small \par}
\end{description}
\end{abstract}

@@

In conventional quantum theory, when we measure a property $\mathbf{O}$
of a physical system whose state is represented by a density matrix~$\rho$,
the probability to obtain the value $\lambda$ of that property is determined by the
Born rule $\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(P_{\lambda}\right)=\Tr\left(\rho P_{\lambda}\right)$,
where $P_{\lambda}$ denotes the projector onto the eigenspace corresponding
to the eigenvalue $\lambda$ of the Hermitian operator
$\mathbf{O}$~\citep{Born1983,peres1995quantum,544199,Jaeger2007}. 
Alternatively, one may say that the state $\rho$ determines the 
probability measure $\mu_{\rho}^{\mathrm{\mathrm{B}}}$ defined as a
function from the set $\events$ of all possible projectors $P$ onto the interval 
$[0,1]$, and satisfies the following probability axioms~
\citep{10.2307/2308516,gleason1957,Redhead1987-REDINA,Maassen2010}: 
\begin{itemize}
\item $\mu_{\rho}^{\mathrm{\mathrm{B}}}(\mathbb{0})=0$. 
\item $\mu_{\rho}^{\mathrm{\mathrm{B}}}(\mathbb{1})=1$. 
\item $\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\mathbb{1}-P\right)=
1-\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(P\right)$. 
\item For a set of mutually orthogonal projectors $\left\{ P_{i}\right\} _{i=1}^{N}$,
$P_iP_j=\delta_{ij}P_i$,
we have 
\begin{equation}
\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\sum_{i=1}^{N}P_{i}\right)=
\sum_{i=1}^{N}\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(P_{i}\right)
\textrm{ .}\label{eq:QuantumProbability-Addition}
\end{equation}
\end{itemize}
Conversely, given any quantum probability measure~$\mu:\events\rightarrow[0,1]$
satisfying the above axioms, Gleason's theorem asserts that there
exists a unique state~$\rho$ such that $\mu=\mu_{\rho}^{\mathrm{\mathrm{B}}}$
in a Hilbert space of dimension $d\geq3$~\citep{gleason1957,Redhead1987-REDINA,peres1995quantum}.

\paragraph{Quantum Interval-valued Probability Measures ({\rm IVPMs})}

@@

Although considering IVPM is new for a quantum theory, people have
already defined classical IVPMs~\citep{JamisonLodwick2004}, which
are essentially the same as quantum IVPMs on a restricted mutually
commuting sub-event space~$\events'\subseteq\events$, where any
subset of $\events$ is called a sub-event space if it satisfies the
follow conditions.
\begin{itemize}
\item $\mathbb{0}\in\events'$.
\item $\mathbb{1}\in\events'$.
\item For any projection $P\in\events'$, we have $\mathbb{1}-P\in\events'$.
\item For a set of mutually orthogonal projections $\left\{ P_{i}\right\} _{i=1}^{N}\subseteq\events'$,
we have $\sum_{i=1}^{N}P_{i}\in\events'$.
\end{itemize}
Quantum IVPM not only stems from classical IVPM but also generalizes
quantum real-valued probability measures which can be expressed a
limiting case when the collection of intervals are all singletons~$\mathscr{I}_{\infty}=\set{\left[x,x\right]}{x\in\left[0,1\right]}$,
where Gleason's theorem is applied. However, it certainly contains
richer phenomena. For example, we will show the Kochen-Specker theorem~\citep{kochenspecker1967,peres1995quantum,Redhead1987-REDINA}
is another limiting case when the collection of intervals are $\mathscr{I}_{\mathrm{D}}=\left\{ \imposs,\necess\right\} $,
where $\imposs=\left[0,0\right]$ and $\necess=\left[1,1\right]$,
in the next paragraph.

@@

The Kochen-Specker theorem asserts that there is no function~$v$
assigning a definite value to all observables such that \begin{subequations}
\label{eq:Kochen-Specker-Rules} 
\begin{eqnarray}
 &  & v\left(\mathbf{O}_{1}\right)+v\left(\mathbf{O}_{2}\right)=v\left(\mathbf{O}_{1}+\mathbf{O}_{2}\right)\textrm{ and}\\
 &  & v\left(\mathbf{O}_{1}\right)v\left(\mathbf{O}_{2}\right)=v\left(\mathbf{O}_{1}\mathbf{O}_{2}\right)
\end{eqnarray}
\end{subequations} if $\mathbf{O}_{1}$ and $\mathbf{O}_{2}$ are
commuting observables. Suppose that there is a function~$v$ satisfying
these two properties, it must send an observable to its eigenvalues
so that the following observables~\citep{Mermin1990Simple,peres1995quantum}:
\begin{equation}
\begin{array}{ccc}
\mathbb{1}\otimes\sigma_{z} & \sigma_{z}\otimes\mathbb{1} & \sigma_{z}\otimes\sigma_{z}\\
\sigma_{x}\otimes\mathbb{1} & \mathbb{1}\otimes\sigma_{x} & \sigma_{x}\otimes\sigma_{x}\\
\sigma_{x}\otimes\sigma_{z} & \sigma_{z}\otimes\sigma_{x} & \sigma_{y}\otimes\sigma_{y}
\end{array}\label{eq:MerminSquare}
\end{equation}
must be sent to $1$ or $-1$. In each row and column, operators commutes,
and each operator is the product of the two others, except in the
third column $\left(\sigma_{z}\otimes\sigma_{z}\right)\left(\sigma_{x}\otimes\sigma_{x}\right)=-\sigma_{y}\otimes\sigma_{y}$.
Therefore, it is clear impossible to find a function~$v$ such that
\begin{equation}
\begin{array}{ccc}
v\left(\mathbb{1}\otimes\sigma_{z}\right) & v\left(\sigma_{z}\otimes\mathbb{1}\right) & v\left(\sigma_{z}\otimes\sigma_{z}\right)\\
v\left(\sigma_{x}\otimes\mathbb{1}\right) & v\left(\mathbb{1}\otimes\sigma_{x}\right) & v\left(\sigma_{x}\otimes\sigma_{x}\right)\\
v\left(\sigma_{x}\otimes\sigma_{z}\right) & v\left(\sigma_{z}\otimes\sigma_{x}\right) & v\left(\sigma_{y}\otimes\sigma_{y}\right)
\end{array}\label{eq:MerminSquare-values}
\end{equation}
satisfying equation~(\ref{eq:Kochen-Specker-Rules}).

@@

Since observables in equation~(\ref{eq:MerminSquare}) are not projections,
if we want to reformulate the above argument in the language of quantum
IVPMs, we first need to find the common eigenbasis of each row and
column, and consider the projections to each eigenvectors~\citep{Kernaghan1994,peres1995quantum}.
Let $\mathcal{T}_{24}$ denote all eigenstates listed in Table~\ref{table:commonBasis}.
The Kochen-Specker theorem could then be proved by showing there is
no function~$v:\events_{24}\rightarrow\left\{ 0,1\right\} $ satisfying
equation~(\ref{eq:Kochen-Specker-Rules}), where 
\begin{equation}
\events_{24}=\set{\sum_{i=1}^{N}\proj{\psi_{i}}}{\begin{gathered}\textrm{mutually orthogonal }\\
\left\{ \ket{\psi_{i}}\right\} _{i=1}^{N}\subseteq\mathcal{T}_{24}
\end{gathered}
}
\end{equation}
is a sub-event space.

@@

Recall we model the idea of the finite precision measurement as IVPM.
When we made finite number of observations, we can use some quantum
IVPMs to represent what we observed. Given a quantum IVPM~$\bar{\mu}$,
the question is whether we can refine precisely enough to determine
the underlying unknown quantum real-valued probability measure, i.e.,
a quantum state. If we can, we say $\bar{\mu}$ has a non-empty core.
More precisely, A \emph{core} of $\bar{\mu}$, $\mathrm{core}\left(\bar{\mu}\right)$,
is the set 
\begin{equation}
\set{\pmeas:\events\rightarrow[0,1]}{\forall P\in\events.~\pmeas\left(P\right)\in\bar{\mu}\left(P\right)}\textrm{ .}
\end{equation}
For example, we can add an one extra interval~$\unknown=\left[0,1\right]$
to $\mathscr{I}_{\mathrm{D}}$, and consider $\mathscr{I}_{1}=\left\{ \imposs,\unknown,\necess\right\} $.
Let 
\begin{equation}
\iota_{1}(x)=\begin{cases}
\necess & \text{, if }x=1\text{ ;}\\
\imposs & \text{, if }x=0\text{ ;}\\
\unknown & \text{, otherwise.}
\end{cases}\label{eq:homomorphism-1}
\end{equation}
Then, $\iota_{1}\circ\mu_{\ket{0}}^{\mathrm{\mathrm{B}}}:\events\rightarrow\mathscr{I}_{1}$
is a quantum IVPM whose core has a unique element~$\mu_{\ket{0}}^{\mathrm{\mathrm{B}}}$,
where $\circ$ denote function composition.

@@

\begin{thm}[Shapley~\citep{Shapley1971,Grabisch2016}]\label{thm:Shapley}Every
classical convex interval-valued probability measure has a non-empty
core, where $\bar{\mu}$ is called convex if 
\begin{equation}
\bar{\mu}\left(P_{0}\vee P_{1}\right)+\bar{\mu}\left(P_{0}P_{1}\right)\subseteq\bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)\label{eq:QuantumInterval-valuedProbability-Convex}
\end{equation}
for all commuting $P_{0},P_{1}\in\events$, and the \emph{disjunction}~$P_{0}\vee P_{1}$
is defined to be $P_{0}+P_{1}-P_{0}P_{1}$~\citep{Griffiths2003}.\end{thm}

>>>>

\documentclass[english,reprint, aps, prl,superscriptaddress, showpacs,
showkeys, longbibliography]{revtex4-1} 
\usepackage{fullpage}
\usepackage[T1,T2A]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{babel}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{mathrsfs}
\usepackage{bbold}
\usepackage{wesa}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage{hyperref}

@@

\author{Andrew J Hanson}
\affiliation{School of Informatics and Computing, Indiana University, Bloomington,
Indiana 47405, USA}

\author{Gerardo Ortiz}
\affiliation{Department of Physics, Indiana University, Bloomington, Indiana 47405,
USA}

\author{Amr Sabry}
\affiliation{School of Informatics and Computing, Indiana University, Bloomington,
Indiana 47405, USA}

\author{Yu-Tsung Tai}
\affiliation{Department of Mathematics, Indiana University, Bloomington, Indiana 47405,
USA}
\affiliation{School of Informatics and Computing, Indiana University, Bloomington,
Indiana 47405, USA}

@@

\begin{abstract}
\begin{description}
\item [{Background}] In a world of bounded resource, any quantum measurement
can be only performed in finite precision.
\item [{Method}] We model the idea of the finite precision measurement
as finite number of intervals, and replace quantum real-valued probability
measures by interval-valued ones.
\item [{Results}] Quantum interval-valued probability measures provide
a unified framework to interpret Gleason's theorem and the Kochen-Specker
theorem as the limiting cases. For the cases between them, we found
some legitimate quantum interval-valued probability measures which
cannot be explained by the Born rule. While restricting on only commuting
projectors, the interval-valued probability measures become classical,
and these non-Born interval-valued probability measures disappear.
\item [{Conclusions}] Our research provides a new insight of finite precision
measurement and non-commutativity when formulating quantum theory.
Also, the Born rule may not be as nature as we used to believe in
the world of finite resource.
\end{description}
\end{abstract}

@@

In conventional quantum theory, when we measure a property $\mathbf{O}$
of a physical system whose state is represented by a density matrix~$\rho$,
the probability to obtain the value $\lambda$ of that property is determined by the
Born rule $\mu_{\rho}^{\mathrm{B}}\left(P_{\lambda}\right)=\Tr\left(\rho P_{\lambda}\right)$,
where $P_{\lambda}$ denotes the projector onto the eigenspace corresponding
to the eigenvalue $\lambda$ of the Hermitian operator
$\mathbf{O}$~\cite{Born1983,peres1995quantum,544199,Jaeger2007}. 
Alternatively, one may say that the state $\rho$ determines the 
probability measure $\mu_{\rho}^{\mathrm{B}}$ defined as a
function from the set $\events$ of all possible projectors $P$ onto the interval 
$[0,1]$, and satisfies the following probability axioms~
\cite{10.2307/2308516,gleason1957,Redhead1987-REDINA,Maassen2010}: 
\begin{itemize}
\item $\mu_{\rho}^{\mathrm{B}}(\mathbb{0})=0$. 
\item $\mu_{\rho}^{\mathrm{B}}(\mathbb{1})=1$. 
\item $\mu_{\rho}^{\mathrm{B}}\left(\mathbb{1}-P\right)=
1-\mu_{\rho}^{\mathrm{B}}\left(P\right)$. 
\item For a set of mutually orthogonal projectors $\left\{ P_{i}\right\} _{i=1}^{N}$,
$P_iP_j=\delta_{ij}P_i$,
we have 
\begin{equation}
\mu_{\rho}^{\mathrm{B}}\left(\sum_{i=1}^{N}P_{i}\right)=
\sum_{i=1}^{N}\mu_{\rho}^{\mathrm{B}}\left(P_{i}\right)
\textrm{ .}\label{eq:QuantumProbability-Addition}
\end{equation}
\end{itemize}
Conversely, given any quantum probability measure~$\mu:\events\rightarrow[0,1]$
satisfying the above axioms, Gleason's theorem asserts that there
exists a unique state~$\rho$ such that $\mu=\mu_{\rho}^{\mathrm{B}}$
in a Hilbert space of dimension $d\geq3$~\cite{gleason1957,Redhead1987-REDINA,peres1995quantum}.

\paragraph{Quantum Interval-valued Probability Measures (\emph{IVPMs})}

@@

Although considering IVPM is new for a quantum theory, people have
already defined classical IVPMs~\cite{JamisonLodwick2004}, which
are essentially the same as quantum IVPMs on a restricted mutually
commuting sub-event space~$\events'\subseteq\events$, where any
subset of $\events$ is called a sub-event space if it satisfies the
follow conditions.
\begin{itemize}
\item $\mathbb{0}\in\events'$.
\item $\mathbb{1}\in\events'$.
\item For any projection $P\in\events'$, we have $\mathbb{1}-P\in\events'$.
\item For a set of mutually orthogonal projections $\left\{ P_{i}\right\} _{i=1}^{N}\subseteq\events'$,
we have $\sum_{i=1}^{N}P_{i}\in\events'$.
\end{itemize}
Quantum IVPM not only stems from classical IVPM but also generalizes
quantum real-valued probability measures which can be expressed a
limiting case when the collection of intervals are all singletons~$\mathscr{I}_{\infty}=\set{\left[x,x\right]}{x\in\left[0,1\right]}$,
where Gleason's theorem is applied. However, it certainly contains
richer phenomena. For example, we will show the Kochen-Specker theorem~\cite{kochenspecker1967,peres1995quantum,Redhead1987-REDINA}
is another limiting case when the collection of intervals are $\mathscr{I}_{\mathrm{D}}=\left\{ \imposs,\necess\right\} $,
where $\imposs=\left[0,0\right]$ and $\necess=\left[1,1\right]$,
in the next paragraph.

@@

The Kochen-Specker theorem asserts that there is no function~$v$
assigning a definite value to all observables such that \begin{subequations}
\label{eq:Kochen-Specker-Rules} 
\begin{eqnarray}
 &  & v\left(\mathbf{O}_{1}\right)+v\left(\mathbf{O}_{2}\right)=v\left(\mathbf{O}_{1}+\mathbf{O}_{2}\right)\textrm{ and}\\
 &  & v\left(\mathbf{O}_{1}\right)v\left(\mathbf{O}_{2}\right)=v\left(\mathbf{O}_{1}\mathbf{O}_{2}\right)
\end{eqnarray}
\end{subequations} if $\mathbf{O}_{1}$ and $\mathbf{O}_{2}$ are
commuting observables. Suppose that there is a function~$v$ satisfying
these two properties, it must send an observable to its eigenvalues
so that the following observables~\cite{Mermin1990Simple,peres1995quantum}:
\begin{equation}
\begin{array}{ccc}
\mathbb{1}\otimes\sigma_{z} & \sigma_{z}\otimes\mathbb{1} & \sigma_{z}\otimes\sigma_{z}\\
\sigma_{x}\otimes\mathbb{1} & \mathbb{1}\otimes\sigma_{x} & \sigma_{x}\otimes\sigma_{x}\\
\sigma_{x}\otimes\sigma_{z} & \sigma_{z}\otimes\sigma_{x} & \sigma_{y}\otimes\sigma_{y}
\end{array}\label{eq:MerminSquare}
\end{equation}
must be sent to $1$ or $-1$. In each row and column, operators commutes,
and each operator is the product of the two others, except in the
third column $\left(\sigma_{z}\otimes\sigma_{z}\right)\left(\sigma_{x}\otimes\sigma_{x}\right)=-\sigma_{y}\otimes\sigma_{y}$.
Therefore, it is clear impossible to find a function~$v$ such that
\begin{equation}
\begin{array}{ccc}
v\left(\mathbb{1}\otimes\sigma_{z}\right) & v\left(\sigma_{z}\otimes\mathbb{1}\right) & v\left(\sigma_{z}\otimes\sigma_{z}\right)\\
v\left(\sigma_{x}\otimes\mathbb{1}\right) & v\left(\mathbb{1}\otimes\sigma_{x}\right) & v\left(\sigma_{x}\otimes\sigma_{x}\right)\\
v\left(\sigma_{x}\otimes\sigma_{z}\right) & v\left(\sigma_{z}\otimes\sigma_{x}\right) & v\left(\sigma_{y}\otimes\sigma_{y}\right)
\end{array}\label{eq:MerminSquare-values}
\end{equation}
satisfying equation~(\ref{eq:Kochen-Specker-Rules}).

@@

Since observables in equation~(\ref{eq:MerminSquare}) are not projections,
if we want to reformulate the above argument in the language of quantum
IVPMs, we first need to find the common eigenbasis of each row and
column, and consider the projections to each eigenvectors~\cite{Kernaghan1994,peres1995quantum}.
Let $\mathcal{T}_{24}$ denote all eigenstates listed in Table~\ref{table:commonBasis}.
The Kochen-Specker theorem could then be proved by showing there is
no function~$v:\events_{24}\rightarrow\left\{ 0,1\right\} $ satisfying
equation~(\ref{eq:Kochen-Specker-Rules}), where 
\begin{equation}
\events_{24}=\set{\sum_{i=1}^{N}\proj{\psi_{i}}}{\begin{gathered}\textrm{mutually orthogonal }\\
\left\{ \ket{\psi_{i}}\right\} _{i=1}^{N}\subseteq\mathcal{T}_{24}
\end{gathered}
}
\end{equation}
is a sub-event space.

@@

Recall we model the idea of the finite precision measurement as IVPM.
When we made finite number of observations, we can use some quantum
IVPMs to represent what we observed. Given a quantum IVPM~$\bar{\mu}$,
the question is whether we can refine precisely enough to determine
the underlying unknown quantum real-valued probability measure, i.e.,
a quantum state. If we can, we say $\bar{\mu}$ has a non-empty core.
More precisely, A \emph{core} of $\bar{\mu}$, $\mathrm{core}\left(\bar{\mu}\right)$,
is the set 
\begin{equation}
\set{\pmeas:\events\rightarrow[0,1]}{\forall P\in\events.~\pmeas\left(P\right)\in\bar{\mu}\left(P\right)}\textrm{ .}
\end{equation}
For example, we can add an one extra interval~$\unknown=\left[0,1\right]$
to $\mathscr{I}_{\mathrm{D}}$, and consider $\mathscr{I}_{1}=\left\{ \imposs,\unknown,\necess\right\} $.
Let 
\begin{equation}
\iota_{1}(x)=\begin{cases}
\necess & \text{, if }x=1\text{ ;}\\
\imposs & \text{, if }x=0\text{ ;}\\
\unknown & \text{, otherwise.}
\end{cases}\label{eq:homomorphism-1}
\end{equation}
Then, $\iota_{1}\circ\mu_{\ket{0}}^{\mathrm{B}}:\events\rightarrow\mathscr{I}_{1}$
is a quantum IVPM whose core has a unique element~$\mu_{\ket{0}}^{\mathrm{B}}$,
where $\circ$ denote function composition.

@@

\begin{thm}[Shapley~\cite{Shapley1971,Grabisch2016}]\label{thm:Shapley}Every
classical convex interval-valued probability measure has a non-empty
core, where $\bar{\mu}$ is called convex if 
\begin{equation}
\bar{\mu}\left(P_{0}\vee P_{1}\right)+\bar{\mu}\left(P_{0}P_{1}\right)\subseteq\bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)\label{eq:QuantumInterval-valuedProbability-Convex}
\end{equation}
for all commuting $P_{0},P_{1}\in\events$, and the \emph{disjunction}~$P_{0}\vee P_{1}$
is defined to be $P_{0}+P_{1}-P_{0}P_{1}$~\cite{Griffiths2003}.\end{thm}
